Fractional Types

### 2019-11-17 fracGC/p.tex
paper

@@

@@

@@

@@

@@

General allocation and de-allocation. No need to keep track of
lifetime. Just make sure location is back to initial value. Safe to
de-allocate because indistinguishable from freshly created. No pointer
equality and aliases though. Concurrent: one thread allocates; gc
thread meets someone else

%%%
\subsection{Examples}

@@

@@

The management of ancilla bits can be teased into two relatively
distinct subproblems:
\begin{itemize}
\item Every allocated ancilla bit must be de-allocated, and 
\item De-allocation of an ancilla bit is only safe if the bit is
  restored to its allocation-time initial value.
\end{itemize}
In this section, we focus on the first subproblem and address the
second in later sections.

@@

@@

When ancilla bits are created they are created with a specific
value. This value is encoded in the circuit. The value of fractional
type represents the ``garbage collector.''  The garbage collector is
specialized to collect a single value, which is the value used to
create the ancilla value. At this point, if the garbage collector
encounters a different value, it doesn't know what to do, and would
need to throw an exception. This situation in \verb|PiFracDyn| is
similar to current solutions (e.g. Quipper) but actually more general
as it allows non-scoped ancilla allocation and de-allocation. Show
examples.

>>>>

Restricting a reversible (classical or quantum) circuit to use no
ancillae is like restricting a Turing machine to use no memory other
than the $n$ bits used to represent the
input~\cite{DBLP:conf/innovations/AaronsonGS17}. As such a restriction
disallows countless computations for trivial reasons, reversible and
quantum models of computation have, since their inception, included
management for scratch storage in the form of ancilla
bits~\cite{Toffoli:1980} with the fundamental restriction that such
bits must be returned to their initial states before being safely
reused or de-allocated.

@@

Beyond computability and efficiency reasons, ancillae have found a
wide range of interesting applications.

@@

\paragraph*{Reversible Branching Operations.}

@@

\paragraph*{Error Correction.}

@@

\paragraph*{Deferring Quantum Measurements.}

@@

@@

%%%
\subsection{A Small Example}

some that ``wants'' ancilla; perhaps fig2b from Dyn. if we want to use
gates with small number of inputs; need to break cccnot to smaller gates

@@

The management of ancilla data can be teased into two relatively
distinct subproblems:
\begin{itemize}
\item Every allocated ancilla bit must be de-allocated, and 
\item De-allocation of an ancilla bit is only safe if the bit is
  restored to its allocation-time initial value.
\end{itemize}
In this section, we focus on the first subproblem and address the
second in later sections.

@@

% of course need de-allocation; keep track of lifetime

% easy: stack but not general

% for general; programmer control or gc

% context of reversible; similar; need allocation; same tradeoffs for
% de-allocation; stack or programmer or gc BUT with caveat. value must
% be restored to its initial value; no information leakage which would
% destroy quantum state

% but reversible also gives you more guarantees; no creation or deletion
% of information means no need to keep track of lifetime of individual
% locations; just need to keep track of ``information''

% General allocation and de-allocation. No need to keep track of
% lifetime. Just make sure location is back to initial value. Safe to
% de-allocate because indistinguishable from freshly created. No pointer
% equality and aliases though. Concurrent: one thread allocates; gc
% thread meets someone else

@@

When ancilla are created they are created with a specific
value. This value is encoded in the circuit. The value of fractional
type represents the ``garbage collector.''  The garbage collector is
specialized to collect a single value, which is the value used to
create the ancilla value. At this point, if the garbage collector
encounters a different value, it doesn't know what to do, and would
need to throw an exception. This situation in \verb|PiFracDyn| is
similar to current solutions (e.g. Quipper) but actually more general
as it allows non-scoped ancilla allocation and de-allocation. Show
examples.
